---
title: "BaggingForests"
author: "Rukh Agha"
date: "2022-07-31"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Bagging and Random Forests

```{r, include=FALSE}
library(ggplot2)
library(cowplot)
library(randomForest)
library(dplyr)

train = read.csv('Training.csv')
test = read.csv('Test.csv')

####Transform membership type from 1 string categorical variable to 10 dummy variables. This drops % variation explained to 0 (actually negative) and significantly increases MSE. Unnecessary for a tree-based model to convert to dummy integers.

# train$Local365 = train$Membership.Type
# train$DayWalkUpPass = train$Membership.Type
# train$Local30 = train$Membership.Type
# train$Explorer = train$Membership.Type
# train$WalkUp = train$Membership.Type
# train$Local31 = train$Membership.Type
# train$Weekender = train$Membership.Type
# train$DayKiosk = train$Membership.Type
# train$UTStudent = train$Membership.Type
# train$Student = train$Membership.Type
# 
# train$Local365 = ifelse(train$Local365=='Local365',1,0)
# train$DayWalkUpPass = ifelse(train$Local365=='24 Hour Walk Up Pass',1,0)
# train$Local30 = ifelse(train$Local365=='Local30',1,0)
# train$Explorer = ifelse(train$Local365=='Explorer',1,0)
# train$WalkUp = ifelse(train$Local365=='Walk Up',1,0)
# train$Local31 = ifelse(train$Local365=='Local 31',1,0)
# train$Weekender = ifelse(train$Local365=='Weekender',1,0)
# train$DayKiosk = ifelse(train$Local365=='24-Hour Kiosk (Austin B-cycle)',1,0)
# train$UTStudent = ifelse(train$Local365=='U.T. Student Membership',1,0)
# train$Student = ifelse(train$Local365=='Student Membership',1,0)
# 
# train = train[-1]
# 
# test$Local365 = test$Membership.Type
# test$DayWalkUpPass = test$Membership.Type
# test$Local30 = test$Membership.Type
# test$Explorer = test$Membership.Type
# test$WalkUp = test$Membership.Type
# test$Local31 = test$Membership.Type
# test$Weekender = test$Membership.Type
# test$DayKiosk = test$Membership.Type
# test$UTStudent = test$Membership.Type
# test$Student = test$Membership.Type
# 
# test$Local365 = ifelse(test$Local365=='Local365',1,0)
# test$DayWalkUpPass = ifelse(test$Local365=='24 Hour Walk Up Pass',1,0)
# test$Local30 = ifelse(test$Local365=='Local30',1,0)
# test$Explorer = ifelse(test$Local365=='Explorer',1,0)
# test$WalkUp = ifelse(test$Local365=='Walk Up',1,0)
# test$Local31 = ifelse(test$Local365=='Local 31',1,0)
# test$Weekender = ifelse(test$Local365=='Weekender',1,0)
# test$DayKiosk = ifelse(test$Local365=='24-Hour Kiosk (Austin B-cycle)',1,0)
# test$UTStudent = ifelse(test$Local365=='U.T. Student Membership',1,0)
# test$Student = ifelse(test$Local365=='Student Membership',1,0)
# 
# test = test[-1]
```

### Bagging - Create random forest with all variables (m=p=19)
```{r echo=FALSE}
set.seed(42)
n = nrow(train)

print('Creating random forest of 20 trees..')
bag_20 = randomForest(TripCount ~., data = train, ntree=20, mtry=19)
print(bag_20)

print('Creating random forest of 50 trees..')
bag_50 = randomForest(TripCount ~., data = train, ntree=50, mtry=19)
print(bag_50)

print('Creating random forest of 200 trees..')
bag_200 = randomForest(TripCount ~., data = train, ntree=200, mtry=19)
print(bag_200)
plot(bag_200)
```

### Bagging with validation set
```{r, echo=FALSE}
#create validation set separated into 2 sets: x inputs for prediction, y values for validation
pred = subset(test, select = -TripCount)
predY = as.double(test$TripCount)

print('Creating random forest of 150 trees..')
bag_val_150 = randomForest(TripCount ~., data = train, xtest = pred, ytest = predY, ntree=150, mtry=19)
print(bag_val_150)
plot(bag_val_150)
```


##random forests
```{r, echo=FALSE}
print('Creating random forest of 50 trees..')
rf_50 = randomForest(TripCount ~., data = train, ntree=50)
print(rf_50)

print('Creating random forest of 150 trees..')
rf_150 = randomForest(TripCount ~., data = train, ntree=150)
print(rf_150)

print('Creating random forest of 300 trees..')
rf_300 = randomForest(TripCount ~., data = train, ntree=300)
print(rf_300)
plot(rf_300)
```


#rf validation
```{r, echo=FALSE}
rf_val_150 = randomForest(TripCount ~., data = train, xtest = pred, ytest = predY, ntree=150)

print(rf_val_150)
plot(rf_val_150)

importance(rf_val_150)
varImpPlot(rf_val_150)

###other work - distance from mccombs
mlat = 30.2840788
mlong = -97.7376486

train = mutate(train, distanceMcCombs = sqrt((Lat-mlat)^2+(Long-mlong)^2))
test = mutate(test, distanceMcCombs = sqrt((Lat-mlat)^2+(Long-mlong)^2))

print(19)
head(test)
str(test)
```






# Appendix

### Bagging without year
```{r echo=FALSE}
train = train[-5]
test = test[-5]

set.seed(7)

print('Creating random forest of 150 trees..')
bag_150 = randomForest(TripCount ~., data = train, ntree=150, mtry=18)
print(bag_150)
plot(bag_150)
```

### Validation set
```{r, echo=FALSE}
#create validation set separated into 2 sets: x inputs for prediction, y values for validation
pred = subset(test, select = -TripCount)
predY = as.double(test$TripCount)

print('Creating random forest of 150 trees..')
bag_val_150 = randomForest(TripCount ~., data = train, xtest = pred, ytest = predY, ntree=150, mtry=18)
print(bag_val_150)
plot(bag_val_150)
```




###Other Work

###
plot(train$Temp, train$TripCount)
plot(train$distance, train$TripCount)
plot(train$lat_distance, train$TripCount)
plot(train$Lat, train$TripCount)

lmdistance = lm(TripCount ~ distance, data = train)
summary(lmdistance)

pred = subset(test, select = -TripCount)
predY = as.double(test$TripCount)
str(pred)
str(predY)

rf2 = randomForest(TripCount ~., data = train, xtest = pred, ytest = predY, ntree=200, mtry=15)
print(rf2)

plot(rf2)

rf3 = randomForest(TripCount ~., data = train, xtest = pred, ytest = predY, ntree=200)
print(rf3)

plot(rf3)